Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-12-20 WSSSPE3/report/WSSSPE3_review.tex
fixed question for D.Gunter lightning talk
missing word in abstract

\begin{abstract}
This technical report records and discusses the Third Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE3). 
%The workshop used an
%alternative submission and peer-review process, which led to a set of papers
%divided across five topic areas: 
The report includes a description of keynote presentation and a set of lightning
talks, and focuses on the discussions, future steps, and future organization for
a set of self-organized working groups.
For each group, the report also includes a point of contact and a landing
page that can be used by those who want to join that group's future activities.
\todo{think about if there are any general conclusions we want to highlight from WSSSPE3}
%%
%The workshop recognized that reliance on scientific software
%is pervasive in all areas of world-leading research today. The workshop
%participants then proceeded to explore different perspectives on the concept of
%sustainability. Key enablers and barriers of sustainable scientific software
%were identified from their experiences. In addition,
%recommendations with new requirements such as software credit files and software
%prize frameworks were outlined for improving practices in sustainable software
%engineering.
%%
%There was also broad consensus that formal
%training in software development or engineering was rare among the
%practitioners. Significant strides need to be made in building a sense of
%community via training in software and technical practices, on increasing their
%size and scope, and on better integrating them directly into graduate education
%programs.
%%
%Finally, journals can define and publish policies to improve reproducibility, whereas
%reviewers can insist that authors provide sufficient information and access to
%data and software to allow them reproduce the results in the paper. Hence a list of
%criteria is compiled for  journals to provide to reviewers so as to make it easier to
%review software submitted for publication as a ``Software Paper.''

@@

\item \textbf{Dan Gunter, Adam Arkin, Rick Stevens, Robert Cottingham, and
Sergei Maslov: \textit{Challenges of a Sustainable Software Platform for
Predictive Biology: Lessons Learned on the KBase Project}} In this talk, Dan
Gunter presented experiences and lessons learned as part of KBase, an open
software and data platform for addressing the grand challenge of systems
biology: predicting and designing biological function. KBase is a unified system
that integrates data and analytical tools for comparative functional genomics of
microbes, plants, and their communities. However, it is also a collaborative
environment for sharing methods and results, and placing those results in the
context of knowledge in the field. Being a large, multi-institutional project,
one of the big challenges is to agree on standard to enable a single,
maintainable system. Working in isolation does not work (anymore) within this
field, and the community in the field also didn't have standards for software
engineering. This is contrasted to computer science research, where these
approaches shorten design cycles, leading to more time for highly rewarded
activities like publishing, performance studies, graduating students, or
protecting ideas before publication. Sustainability is not of big concern there,
compared to these. \katznote{prev sentence is unclear} Instead of this more
traditional approach, KBase uses the following for large, distributed teams.
After picking projects and team members, 4-5 teams work on projects about two
weeks before a 1-week evaluation by the executive committee, and after which new
teams and projects might be chosen to restart the ``agile'' development cycle.
This process is intentionally open and documented.

>>>>

\begin{abstract}
This technical report records and discusses the Third Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE3). 
%The workshop used an
%alternative submission and peer-review process, which led to a set of papers
%divided across five topic areas: 
The report includes a description of the keynote presentation and a set of lightning
talks, and focuses on the discussions, future steps, and future organization for
a set of self-organized working groups.
For each group, the report also includes a point of contact and a landing
page that can be used by those who want to join that group's future activities.
\todo{think about if there are any general conclusions we want to highlight from WSSSPE3}
%%
%The workshop recognized that reliance on scientific software
%is pervasive in all areas of world-leading research today. The workshop
%participants then proceeded to explore different perspectives on the concept of
%sustainability. Key enablers and barriers of sustainable scientific software
%were identified from their experiences. In addition,
%recommendations with new requirements such as software credit files and software
%prize frameworks were outlined for improving practices in sustainable software
%engineering.
%%
%There was also broad consensus that formal
%training in software development or engineering was rare among the
%practitioners. Significant strides need to be made in building a sense of
%community via training in software and technical practices, on increasing their
%size and scope, and on better integrating them directly into graduate education
%programs.
%%
%Finally, journals can define and publish policies to improve reproducibility, whereas
%reviewers can insist that authors provide sufficient information and access to
%data and software to allow them reproduce the results in the paper. Hence a list of
%criteria is compiled for  journals to provide to reviewers so as to make it easier to
%review software submitted for publication as a ``Software Paper.''

@@

\item \textbf{Dan Gunter, Adam Arkin, Rick Stevens, Robert Cottingham, and
Sergei Maslov: \textit{Challenges of a Sustainable Software Platform for
Predictive Biology: Lessons Learned on the KBase Project}} In this talk, Dan
Gunter presented experiences and lessons learned as part of KBase, an open
software and data platform for addressing the grand challenge of systems
biology: predicting and designing biological function. KBase is a unified system
that integrates data and analytical tools for comparative functional genomics of
microbes, plants, and their communities. However, it is also a collaborative
environment for sharing methods and results, and placing those results in the
context of knowledge in the field. Being a large, multi-institutional project,
one of the big challenges is to agree on standard to enable a single,
maintainable system. Working in isolation does not work (anymore) within this
field, and the community in the field also didn't have standards for software
engineering. This is contrasted to computer science research, where these
approaches shorten design cycles, leading to more time for highly rewarded
activities like publishing, performance studies, graduating students, or
protecting ideas before publication. Instead of this more
traditional approach, KBase uses a variation on the ``Scrum'' methodology.
After picking projects and team members, 4-5 teams work on projects about two
weeks before a 1-week evaluation by the executive committee, and after which new
teams and projects might be chosen to restart the ``agile'' development cycle.
This process is intentionally open and documented.
